### 2022-08-21 chapters/chapter6.tex
update the final version

\chapter{Interpretability of DRL}
\label{ch:drl_interp}
\section{Introduction}
In previous chapters, we have discussed several methods to improve sample efficiency in deep reinforcement learning (DRL), especially in complex robot manipulation tasks with continuous control. However, all of the above experiments are carried out in a simulated environment. Training becomes much more problematic when applying DRL to real-world applications. For example, in a robot manipulation task, which requires an agent to pick up a cube and place it in a target position, computer vision techniques, such as object detection, pose estimation, or semantic segmentation, are used to estimate the position and orientation of the cube. Then, at each timestep, the current information of the cube and the proprioceptive data of the robot arm are provided to the RL agent to select an action.

To reduce \td{real-world training times}, a number of end-to-end approaches~\cite{gu2017deep,levine2016end,zhu2017target} have been proposed in recent years, which enable the agent to infer the policy according to raw observations (e.g., images) directly. However, such an end-to-end training setup would require the agent to collect samples for a long time in a real-world environment to update the policy network, which may increase the risk of robot/hardware failure and thus elevate the cost of training significantly. An emerging solution is to have the agent trained in a simulated environment before transferring the policy to the real world. In this period, the agent also needs to overcome the ``reality gap"~\cite{jakobi1995noise} between the simulated environment and the real-world environment, and domain randomisation (DR)~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain} is a widely used approach for this purpose. When using DR in training, the elements of the simulated environment are varied to some extent, from the visual appearance to the environment dynamics (see Figure~\ref{fig:dr_example}), in order to improve the policy's robustness to various changes in the environment.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/domain_random.png}
  \caption{Examples of domain randomisation (visual appearances) in the simulated Fetch (top) and Jaco (bottom) robots environments.}
  \label{fig:dr_example}
\end{figure}

However, such an end-to-end training paradigm, which could involve inferring actions based on image inputs, makes the theoretical analysis of both the environment and the agents that act within it practically infeasible. Specifically, deep neural networks (NNs) trained with DRL form a direct ``black box" mapping from observation to action, presenting the challenge of understanding learned control strategies. If we would like to deploy such policies---particularly on robots in the real world---we would also want to be able to \emph{explain} them~\cite{arrieta2020explainable}. An interpretation of the inference of a model can be used not only as a way of explaining the behaviour of the model, but can also be used to characterise other properties, such as safety, fairness, and reliability ~\cite{doshi2017towards}. Although there are many existing approaches to explain NNs~\cite{guidotti2018survey}, each of these methods is subjective to varying degrees.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/strategies.png}
  \caption{The range of strategies an agent may learn to use when trained to reach a red target, which we split into three components. Firstly, the agent must visually localise the target---which can be accomplished by detecting red, spherical objects, or simply anything with a significant red component. Secondly, the agent can use vision and/or proprioception to guide its arm to the target. Finally, the agent may accomplish the task with varying levels of robustness to changes in the rest of the visual scene. Using a range of analyses, we show that agents trained to accomplish the same task learn different subsets of these strategies, depending on their training conditions.}
  \label{fig:strategies}
\end{figure}

In this chapter, we aim to understand \td{how DRL agents make decision in the environment} and understand how agents trained with DR behave differently from agents trained without. To this end, we train DRL agents under a range of different settings and use relative differences to better interpret their policies. In contrast to most prior work examining DRL policies~\cite{atrey2020exploratory,greydanus2018visualizing, olson2021counterfactual,puri2020explain,rupprecht2020finding,such2018atari,zahavy2016graying}, we not only provide explanations for how policies react to parts of states/individual states, but use our array of experiments to characterise their \emph{global} strategies (see Figure~\ref{fig:strategies}). \td{Through these experiments and analysis, we are able to address following questions:
\begin{itemize}
    \item What is the role of DR on layer learning?
    \item To what extent is visual inputs rather than proprioceptive inputs used by the agent?
    \item How robust are agents to unseen scenarios?
    \item How redundant is the network in its representation? 
\end{itemize}
}
%For instance, we examine the relative importance of different input modalities, and whether agents need their memory to complete tasks.

In particular, we train different robots with vision-based inputs to perform the task of reaching for a (red) target, and then use different techniques to understand what features of the environment the policies react to. Under certain training settings, the agent may localise the target by simply looking for anything red, and it is only under more challenging training conditions that the agent utilises both colour and shape. Clearly, the latter strategy better represents what we want the agent to learn, while the former is merely a ``shortcut'' \cite{geirhos2020shortcut}.

To be noticed, the structure in this chapter is organised in a different way from the previous ones. In Section~\ref{ch6:sec_interp}, we first introduce the interpretability techniques used in this work. In Section~\ref{ch6:exp}, we detail our customised simulated environments, the network structure of DRL agents, training details, and the settings of test scenarios. Subsequently, in Section~\ref{ch6:results}, we conduct an exhaustive analysis on the trained models using a broad suite of interpretability techniques. Finally, in Section~\ref{ch6:discussion}, we discuss our findings and provide recommendations for applying interpretability techniques to DRL agents.

\section{Interpretability Techniques}
Neural network (NN) has achieved great success in various fields, benefiting from its powerful feature representation capabilities. However, compared with hand-crafted feature descriptors, NN lacks interpretability, and it treats the training as black boxes. Therefore, it is important that we are able to explain and understand what the NN has learned before deploying it in real-world applications, where safety and reliability are crucial. In this work, we focus on understanding the NN in a DRL training paradigm. To do so, various interpretability methods have been considered, which include saliency maps, unit ablations, layer re-initialisation and recurrent ablations. 
\label{ch6:sec_interp}
\subsection{Saliency Maps}
\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_baseline}
    \subcaption{Occ. (Gaussian)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_jaco_large.png}
    \subcaption{Avg. $\mathbf{x}^{base}$}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_average_map.png}
    \subcaption{Occ. (avg. $\mathbf{x}^{base}$)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_mask.png}
    \subcaption{Avg. $\mathbf{x}^{base}$ mask}
  \end{subfigure}
  \caption{Occlusion-based saliency map method (a) using Gaussian mask (c) our improved method using an average image (b, d). Occlusion-based method using Gaussian blur masks (a) shows extra artifacts. Our improved saliency map method highlights important regions in the result by using average baseline image created from plentiful trajectories (c), (d) shows the usage of the average baseline image.}
  \label{fig:saliency_baseline}
\end{figure}

Saliency maps are widely used to understand and explain the output of NNs. Similarly to previous work on understanding DRL agents~\cite{greydanus2018visualizing}, we also use an occlusion-based approach for interpreting. The occlusion-based method~\cite{zeiler2014visualizing} generally adopts a gray mask to scan over the entire input image, and then measures the variation of the output in each masked region to identify which areas have significant influences on the output of the NN. The variation of the performance is also called saliency value, and it can be expressed as:
\begin{align}
S_{m, n} = \lVert F(\mathbf{x}) - F(\mathbf{x}_{m, n}^{occ}) \rVert_2,
\end{align}
where $(m, n)$ is the location on the image, $F(\mathbf{x})$ is the output of the original input, and $F(\mathbf{x}_{m, n}^{occ})$ is the output of the input with an occlusion mask at location $(m, n)$. In the work of~\cite{greydanus2018visualizing}, which aims to understand DRL agents in Atari games, the gray mask is replaced with a Gaussian mask. Because the agent only receives gray scale images in the training and testing. Thus, the gray mask can be recognised as a part of the gray object, and the Gaussian blur mask can introduce spatial uncertainty to find informative regions.

However, the agent receives an RGB image as the input in our task. As a result, the occlusion-based method using Gaussian blur masks cannot highlight any important regions (see Figure~\ref{fig:saliency_baseline}a), and thus fails to provide any valuable information to understand the agent. Atrey \textit{et al.}~\cite{atrey2020exploratory} identify this issue of applying saliency methods to DRL agents as modifying observations in a manner that is incongruent with the true environment's state and generative process, and proposed intervening directly in the environment to employ counterfactual methods. However, in general, this level of control over the environment is not easy. In this work, we are more interested in task-specific objects in the environment, such as the robot arm and the red ball. To avoid interference from the occlusion-based masks on unimportant areas, we first average the frames of approximately 1,000 trajectories to get a baseline image (see Figure~\ref{fig:saliency_baseline} (b)). Then, we use this baseline image to generate the mask in the corresponding locations (see Figure~\ref{fig:saliency_baseline}d). Finally, we can get an improved saliency map, which highlights the informative regions in the input image  (see Figure~\ref{fig:saliency_baseline}c).
%Inspired by the work which use reference images for the calculation of interpreta

Another recent study introduced a more robust DRL-specific saliency method~\cite{puri2020explain}, but it is only applicable to agents which learn a state-action value function over discrete action spaces.
%\subsection{Statistical and Structural Weight Characterisations}
\subsection{Unit Ablations}
Unit ablations evaluate the importance of a single filter in the convolutional layer by masking it to zero and examining the variations in the performance. In general, an important filter can lead to a significant performance drop. While more sophisticated unit ablations have been applied in DRL~\cite{meyes2020you}, {this has only been in the context of a control task with a low-dimensional symbolic state space}.
% Layer initialisation
\subsection{Layer Re-initialisation}
Layer re-initialisation~\cite{zhang2019all} is an extension of unit ablations, which is proposed to test the re-initialisation robustness of the trained neural network. The primary idea is to replace the parameters of a single fully trained layer with the parameters in the previous training epochs, and observe the effect on the performance of the network. In this work, the parameters of each layer can be expressed as: $\{\theta_{1}, \theta_{2}, ..., \theta_{L}\}$, where $L$ is the number of layers of the network, and $\theta_{l}^{t}$ represents the parameters of the $l^\text{th}$ layer in the training epoch $t\in [1,T]$. For example, when training is finished, to evaluate the robustness of a single layer $l$ at training epoch $t$, we can re-initialise the parameters in the layer by: $\theta_{l}^{T} \leftarrow \theta_{l}^{t}$. Furthermore, we can also measure the changes of parameters in the $\ell_{\infty}$ and $\ell_{2}$-norm. Layer re-initialisation allows us to identify the critical layers in the network structure, and these layers are greatly affected by the parameters re-initialisation. Furthermore, with a similar capacity of network, we can also investigate the task complexity by using re-initialisation robustness.

\subsection{Recurrent Ablation}
Recurrent ablation is used to evaluate if the recurrent layer (e.g., long short-term memory (LSTM)) is utilised in a network structure. To do so, we set the hidden state to a constant value. If the performance of the agent is decreased to some extent, it means the recurrent layer helps to perform the task, otherwise, the recurrent layer is redundant in this structure. While it is possible to investigate the hidden states of DRL agents over time, visualising and inspecting a high-dimensional time series can be difficult, even for domain experts~\cite{jaunet2020drlviz}.

@@

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_vol_resize.png}
    \subcaption{Fetch enviroment}
    \label{fig:fetch_volumn}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_view.png}
    \subcaption{Fetch visual input}
    \label{fig:fetch_view}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_vol_resize.png}
    \subcaption{Jaco environment}
    \label{fig:jaco_volumn}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_top.png}
    \subcaption{Jaco visual input}
    \label{fig:jaco_view}
  \end{subfigure}
  \caption{(a) Fetch environment, (b) an example of Fetch visual input, (c) Jaco environment, and (d) an example of Jaco visual input. \td{The blue cuboids indicate the sampling space of the target objects.}}
  \label{fig:robot_setups}
\end{figure}

The Fetch robot environment includes a 7 degrees-of-freedom (DOF) robot arm, and it is modified based on the model provided in OpenAI Gym~\cite{plappert2018multi}. In our Fetch environment, we revise the original proprioceptive input, where we delete the velocity and position of the gripper and the position of the target. It now only contains the position and velocity of 7 joints of the robot arm. In order to acquire visual inputs, we also add an extra RGB camera in front of the robot arm (see Figure~\ref{fig:fetch_view}). The control scheme remains the same, which is the relative position of the gripper in x, y and z axis. The target object is placed on the table. During training, the position of the target object is randomly sampled in the 2D space (see blue cuboid in Figure~\ref{fig:fetch_volumn}) of the table surface. During testing, 80 fixed uniformly distributed positions are used to evaluate the performance of the agent.

The Jaco robot environment includes a 6 DoF robot arm (the gripper is disabled). The proprioceptive input contains the position and velocity of 6 joints. An RGB camera is placed on the top of the robot arm to provide visual inputs (see Figure~\ref{fig:jaco_view}). The action space is the 6 joint velocities of the robot arm. During training, the position of the target object is randomly sampled in a 3D space (see blue cuboid in Figure~\ref{fig:jaco_volumn}). During testing, 250 uniformly distributed fixed positions are used to evaluate the performance of the agent.

@@

\subsection{Training Configurations}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.85\linewidth]{figures/chapter6/axes.png}
  \caption{The 8 different experimental conditions: 2 robots (Fetch vs. Jaco), 2 input modalities (vision vs. vision + proprioception), 2 visual conditions (DR vs. no DR).}
  \label{fig:axes}
\end{figure}
We setup 8 different configurations to cover different possible training strategies and provide a comprehensive comparison. There include different combinations over three primitives: 1) using Fetch robot arm or Jaco robot arm, 2) using visual input only or visual input + proprioceptive input, and 3) training with or without domain randomisation (DR) (see Figure~\ref{fig:axes}).

@@

For real-world deployment, one would want to use as many sensors as it is feasible to use. However, we are interested in how the absence of specific input modalities can affect learning, as this changes how the agent is able to extract information from the environment. For example, while one might expect robots equipped with proprioceptive sensors to use these alone for pose estimation, agents can also learn to utilise vision, if extracting a pose from images is not too difficult.

Finally, analogously to how \emph{data augmentation}~\cite{lecun1998gradient,shorten2019survey} can be used to improve the generalisation of models in supervised learning settings, \emph{domain randomisation} (Figure \ref{fig:dr_example}) is a common training paradigm in ``sim2real'' approaches for learning real-world robots control~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain}. In DR, various properties of the simulation are varied, altering anything from the positions or dynamical properties of objects to their visual appearance, resulting in an expansion of the training domain. We therefore test how the agents trained with or without DR differ. Supporting previous results, we show that agents trained with DR are more robust to out-of-distribution (OoD) perturbations. In our experiments, we use a standard visual DR setup. At each timestep, we randomise all of the components in the environment by changing/adding different textures, colour gradients, RGB colours, and noise, except for the target object. More details about the DR textures are described in Table~\ref{tbl:random_summary} and visualised in Figure~\ref{fig:random_vis}.

%\subsection{Domain Randomisation}
\begin{table}[h!]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|c|c}
    \toprule
    No. & Option & Description \\
    \midrule
   1 & checkerboard & Render checkerboard pattern with two random RGB colours\\
   2 & gradient & Render gradient in vertical/horizontal direction with two random RGB colours\\
   3 & flat\_rgb & Render single random RGB colour\\
   4 & noise & Render artificial noise with two random RGB colours\\
    \bottomrule
  \end{tabular}}
   \caption{DR textures used during training. At every environment timestep, for each component (e.g., robot arm, skybox, table; the Jaco robot has 13 components, and the Fetch robot has 19 components), the appearance of the component is rendered using a sequential sampling process. In the first step of the sampling process, one of each of the 4 possible choices is made with the same probability, then the subsequent appearance of each component is determined by a second draw, described in the rightmost column of the table. Each RGB channel is drawn uniformly from $\{0 \ldots 255\}$. The details of each option are described in Table~\ref{tbl:texture_expression}.}
\label{tbl:random_summary}
\end{table}

>>>>

\chapter{Analysing DRL Agents Trained with Domain Randomisation}
\label{ch:drl_interp}
\section{Introduction}
% In previous chapters, we have discussed several methods to improve sample efficiency in deep reinforcement learning (DRL), especially in complex robot manipulation tasks with continuous, rather than discrete, control. 
In this thesis, we first provide a comprehensive analysis on the trained DRL agents, which aims to understand features and policies that the agents learned in the environment. Specifically, proximal policy optimisation (PPO)~\cite{schulman2017proximal} is selected as the DRL algorithm to train the model, because it is widely used in a variety of scenarios~\cite{berner2019dota,rajeswaran2017learning,vinyals2019grandmaster}, we believe that this analysis can provide more insight to the community. In the following chapters, we will further improve the sample efficiency of PPO algorithm with different ways in the tasks with sparse or delayed rewards. Recently, most DRL experiments are performed in a simulated environment. Training becomes much more problematic when applying DRL to real-world applications. For example, in a robot manipulation task, which requires an agent to pick up a cube and place it in a target position, computer vision techniques, such as object detection, pose estimation, or semantic segmentation, are used to estimate the position and orientation of the cube. Then, at each timestep, the current information of the cube and the proprioceptive data of the robot arm are provided to the RL agent to select an action.

%To reduce \td{real-world training times}, 
A number of end-to-end approaches~\cite{gu2017deep,levine2016end,zhu2017target} have been proposed in recent years, which enable the agent to infer the policy according to raw observations (e.g., images) directly. However, such an end-to-end training setup would require the agent to collect samples for a long time in a real-world environment to update the policy network, which may increase the risk of robot/hardware failure and thus elevate the cost of training significantly. An emerging solution is to have the agent trained in a simulated environment before transferring the policy to the real world. In this period, the agent also needs to overcome the ``reality gap"~\cite{jakobi1995noise} between the simulated environment and the real-world environment, and domain randomisation (DR)~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain} is a widely used approach for this purpose. When using DR in training, the elements of the simulated environment are varied to some extent, from the visual appearance to the environment dynamics (see Figure~\ref{fig:dr_example}), in order to improve the policy's robustness to various changes in the environment.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/domain_random.png}
  \caption[Examples of visual domain randomisation in the simulated robot environments.]{Examples of domain randomisation (visual appearances) in the simulated Fetch (top) and Jaco (bottom) robots environments.}
  \label{fig:dr_example}
\end{figure}

However, such an end-to-end training paradigm, which could involve inferring actions based on image inputs, makes the theoretical analysis of both the environment and the agents that act within it practically infeasible. Specifically, deep neural networks (NNs) trained with DRL form a direct ``black box" mapping from observation to action, presenting the challenge of understanding learned control strategies. If we would like to deploy such policies---particularly on robots in the real world---we would also want to be able to \emph{explain} them~\cite{arrieta2020explainable}. An interpretation of the inference of a model can be used not only as a way of explaining the behaviour of the model, but can also be used to characterise other properties, such as safety, fairness, and reliability ~\cite{doshi2017towards}. Although there are many existing approaches to explain NNs~\cite{guidotti2018survey}, each of these methods is subjective to varying degrees.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{figures/chapter6/strategies.png}
  \caption[The range of strategies an agent may learn to use when trained to reach a red target.]{The range of strategies an agent may learn to use when trained to reach a red target, which we split into three components. Firstly, the agent must visually localise the target---which can be accomplished by detecting red, spherical objects, or simply anything with a significant red component. Secondly, the agent can use vision and/or proprioception to guide its arm to the target. Finally, the agent may accomplish the task with varying levels of robustness to changes in the rest of the visual scene. Using a range of analyses, we show that agents trained to accomplish the same task learn different subsets of these strategies, depending on their training conditions. [This diagram was created by Dr.Kai Arulkumaran and Tamara Gerbert~\cite{dai2022analysing}.]}
  \label{fig:strategies}
\end{figure}

In this chapter, we aim to understand how DRL agents make decision in the environment and understand how agents trained with visual DR behave differently from agents trained without. To this end, we train DRL agents under a range of different settings and use relative differences to better interpret their policies. In contrast to most prior work examining DRL policies~\cite{atrey2020exploratory,greydanus2018visualizing, olson2021counterfactual,puri2020explain,rupprecht2020finding,such2018atari,zahavy2016graying}, we not only provide explanations for how policies react to parts of states/individual states, but use our array of experiments to characterise their \emph{global} strategies (see Figure~\ref{fig:strategies}). Through these experiments and analysis, we are able to provide some insight into the following questions:
\begin{itemize}
    \item How robust are agents to unseen scenarios?
    \item To what extent are visual inputs rather than proprioceptive inputs used by the agent?
    \item How redundant is the network in its representation? 
    \item What is rate of learning in different layers under different training configurations?
\end{itemize}
%For instance, we examine the relative importance of different input modalities, and whether agents need their memory to complete tasks.

In particular, we train different robots with vision-based inputs to perform the task of reaching for a (red) target, and then use different techniques to understand what features of the environment the policies react to. \textit{Unlike} many experiments reported in the literature~\cite{andrychowicz2017hindsight,fang2018dher,fang2019curriculum}, the agent does not have direct access to the coordinates of the ball as a state. Under certain training settings, the agent may localise the target by simply looking for anything red, and it is only under more challenging training conditions that the agent utilises both colour and shape. Clearly, the latter strategy better represents what we want a real-world agent to learn, while the former is merely a ``shortcut'' \cite{geirhos2020shortcut}.

The structure in this chapter is organised in a different way from the previous ones. In Section~\ref{ch6:sec_interp}, we first introduce the interpretability techniques used in this work. In Section~\ref{ch6:exp}, we detail our customised simulated environments, the network structure of DRL agents, training details, and the settings of test scenarios. Subsequently, in Section~\ref{ch6:results}, we conduct an exhaustive analysis on the trained models using a broad suite of interpretability techniques. Finally, in Section~\ref{ch6:discussion}, we discuss our findings and provide recommendations for applying interpretability techniques to DRL agents.

\section{Interpretability Techniques}
Neural networks (NNs) have achieved great success when used with reinforcement learning. However, compared to hand-crafted feature descriptors, NNs lack interpretability, and the networks are often treated as black boxes. Therefore, it is important that we are able to explain and understand what the NN has learned before deploying it in real-world applications, where safety and reliability are crucial. In this work, we focus on understanding the NN in a DRL training paradigm. To do so, various interpretability methods have been considered, which include saliency maps, unit ablations, layer re-initialisation and recurrent ablations. 
\label{ch6:sec_interp}
\subsection{Saliency Maps}
\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_baseline}
    \subcaption{Occ. (Gaussian)}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_jaco_large.png}
    \subcaption{Avg. $\mathbf{x}^{base}$}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/average_map_mask.png}
    \subcaption{Avg. $\mathbf{x}^{base}$ mask}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/occ_jaco_average_map.png}
    \subcaption{Occ. (avg. $\mathbf{x}^{base}$)}
  \end{subfigure}
  \caption[Visualisation of different occlusion-based saliency map methods.]{Occlusion-based saliency map methods: (a) using Gaussian blur mask and (d) using an average image (b, c). Occlusion-based method using Gaussian blur masks (a) shows extra artifacts. Our improved saliency map method (d) highlights important regions in the result by using average baseline image created from plentiful trajectories (b). (c) shows one instance step that using the average baseline image as the occlusion in the original input. The occlusion region is highlighted with a blue outline.}
  \label{fig:saliency_baseline}
\end{figure}

Saliency maps are widely used to understand and explain the output of NNs. Similarly to previous work on understanding DRL agents~\cite{greydanus2018visualizing}, we also use an occlusion-based approach for interpreting. The occlusion-based method~\cite{zeiler2014visualizing} generally adopts a gray mask to scan the entire input image and then measures the variation of the output in each masked region to identify which areas have significant influences on the output of the NN. The variation of the performance is also called the saliency value, and it can be expressed as:
\begin{align}
S_{m, n} = \lVert F(\mathbf{x}) - F(\mathbf{x}_{m, n}^{occ}) \rVert_2,
\end{align}
where $(m, n)$ is the location on the image $\mathbf{x}$, $F(\mathbf{x})$ is the output of the original input, and $F(\mathbf{x}_{m, n}^{occ})$ is the output of the input with an occlusion mask at location $(m, n)$. In the work of~\cite{greydanus2018visualizing}, which aims to understand DRL agents in Atari games, the gray mask is replaced with a Gaussian blur mask. Because the agent only receives gray scale images in the training and testing. Thus, the gray mask can be recognised as a part of the gray object, and the Gaussian blur mask can introduce spatial uncertainty to find informative regions.

However, the agent receives an RGB image as an input in our task. As a result, the occlusion-based method using Gaussian blur masks cannot highlight any important regions (see Figure~\ref{fig:saliency_baseline}a), and thus fails to provide any valuable information to understand the agent. Atrey \textit{et al.}~\cite{atrey2020exploratory} identify this issue of applying saliency methods to DRL agents as modifying observations in a manner that is incongruent with the true environment's state and generative process, and proposed intervening directly in the environment to employ counterfactual methods. However, in general, this level of control over the environment is not easy. In this work, we are more interested in task-specific objects in the environment, such as the robot arm and the red ball. To avoid interference from the occlusion-based masks on unimportant areas, we first sample 1,000 input frames with the trained policy from the environments (the policy is not updated during this process). Next, these 1000 frames are averaged to produce a baseline image (see Figure~\ref{fig:saliency_baseline}b). Then, we use this baseline image to generate the occlusion in the corresponding locations (see Figure~\ref{fig:saliency_baseline}c) in the input frame. Finally, we can measure the performance change and get an improved saliency map, which highlights the informative regions in the input image  (see Figure~\ref{fig:saliency_baseline}d).
%Inspired by the work which use reference images for the calculation of interpreta

Another recent study introduced a more robust DRL-specific saliency method~\cite{puri2020explain}, but it is only applicable to agents which learn a state-action value function over discrete action spaces.
%\subsection{Statistical and Structural Weight Characterisations}
\subsection{Channel-wise Unit Ablations}
Unit ablations evaluate the importance of a single unit in the convolutional layer by masking it to zero and examining the variations in the performance. In general, an important filter can lead to a significant performance drop. If a unit is set to zero and it has no effect on the ability to achieve a task, then it means one of two things: either the unit is redundant and performs no useful function in the network, or the function of the network is well-distributed amongst multiple neurons. While more sophisticated unit ablations have been applied in DRL~\cite{meyes2020you}, {this has only been in the context of a control task with a low-dimensional symbolic state space}. We further provide a Python style code snippet below to demonstrate the process of unit ablations.
\begin{minted}[frame=lines, mathescape, linenos]{python}
# Python code of unit ablations
import torch
for i in range(N_c): # N_c is the number of output channels
    net.load_state_dict(torch.load(ckpt_path))
    net.conv1.weight[i, :, :, :] = 0
    net.conv1.bias[i] = 0
    success_rate = eval(net)
\end{minted}
% Layer initialisation
\subsection{Layer Re-initialisation}
Layer re-initialisation~\cite{zhang2019all} is an extension of unit ablations, which is proposed to test the re-initialisation robustness of the trained neural network. The primary idea is to replace the parameters of a single fully trained layer with the parameters in the previous training epochs, and observe the effect on the performance of the network. In this work, the parameters of each layer can be expressed as: $\{\theta_{1}, \theta_{2}, ..., \theta_{L}\}$, where $L$ is the number of layers in the network, and $\theta_{l}^{t}$ represents the parameters of the $l^\text{th}$ layer in the training epoch $t\in [1,T]$. For example, when training is finished, to evaluate the robustness of a single layer $l$ at training epoch $t$, we can re-initialise the parameters in the layer by: $\theta_{l}^{T} \leftarrow \theta_{l}^{t}$. Furthermore, we can also measure the changes of parameters in the $\ell_{\infty}$ and $\ell_{2}$-norm. Layer re-initialisation allows us to identify the critical layers in the network structure, and these layers are greatly affected by the parameters re-initialisation. Furthermore, with a similar capacity of network, we can also investigate the task complexity by using re-initialisation robustness.

\subsection{Recurrent Ablation}
Recurrent ablation is used to evaluate if the recurrent layer (e.g., long short-term memory (LSTM)) is utilised in a network structure. To do so, we set the hidden and cell states to constant values. If the performance of the agent is decreased to some extent, it means that the recurrent layer helps to perform the task, otherwise, the recurrent layer is redundant in this structure. While it is possible to investigate the hidden states of DRL agents over time, visualising and inspecting a high-dimensional time series can be difficult, even for domain experts~\cite{jaunet2020drlviz}. More details and explanations are in Section~\ref{sec:detail_reucrrent_ablation}.

@@

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_vol_resize.png}
    \subcaption{Fetch enviroment}
    \label{fig:fetch_volumn}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/fetch_view.png}
    \subcaption{Fetch visual input}
    \label{fig:fetch_view}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_vol_resize.png}
    \subcaption{Jaco environment}
    \label{fig:jaco_volumn}
  \end{subfigure}
  \begin{subfigure}{0.24\columnwidth}
    \includegraphics[width=\linewidth]{figures/chapter6/jaco_top.png}
    \subcaption{Jaco visual input}
    \label{fig:jaco_view}
  \end{subfigure}
  \caption[Examples of Fetch and Jaco robot manipulation environments.]{(a) Fetch environment, (b) an example of Fetch visual input, (c) Jaco environment, and (d) an example of Jaco visual input. The blue cuboids indicate the sampling space of the target objects.}
  \label{fig:robot_setups}
\end{figure}

The Fetch robot environment includes a 7 degrees-of-freedom (DOF) robot arm, and it is modified based on the model provided in OpenAI Gym~\cite{plappert2018multi}. In our Fetch environment, we revise the original proprioceptive input, where we delete the velocity and position of the gripper and the position of the target. It now only contains the position and velocity of 7 joints of the robot arm. In order to acquire visual inputs, we also add an extra RGB camera in front of the robot arm (see Figure~\ref{fig:fetch_view}). The control scheme remains the same, which is the relative position of the gripper in $x$, $y$ and $z$ axis. The target object is placed on the table. During training, the position of the target object is randomly sampled in the 2D space (see the blue cuboid in Figure~\ref{fig:fetch_volumn}) of the table surface. During testing, 80 fixed uniformly distributed positions are used to evaluate the performance of the agent.

The Jaco robot environment includes a 6 DoF robot arm (the gripper is disabled). The proprioceptive input contains the position and velocity of 6 joints. An RGB camera is placed on the top of the robot arm to provide visual inputs (see Figure~\ref{fig:jaco_view}). The action space is the 6 joint velocities of the robot arm. During training, the position of the target object is randomly sampled in a 3D space (see the blue cuboid in Figure~\ref{fig:jaco_volumn}). During testing, 250 uniformly distributed fixed positions are used to evaluate the performance of the agent.

@@

\subsection{Training Configurations}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.85\linewidth]{figures/chapter6/axes.png}
  \caption[Visualisation of 8 different experimental conditions.]{The 8 different experimental conditions: 2 robots (Fetch vs. Jaco), 2 input modalities (vision vs. vision + proprioception), 2 visual conditions (DR vs. no DR). [This diagram was created by Dr.Kai Arulkumaran and Tamara Gerbert~\cite{dai2022analysing}.]}
  \label{fig:axes}
\end{figure}
We setup 8 different configurations to cover different possible training strategies and provide a comprehensive comparison. There are different combinations over three primitives: 1) using Fetch robot arm or Jaco robot arm, 2) using visual input only or visual input + proprioceptive input, and 3) training with or without domain randomisation (DR) (see Figure~\ref{fig:axes}).

@@

For a real-world deployment, one would want to use as many sensors as feasible. However, we are interested in how the absence of specific input modalities can affect learning, as this changes the way the agent is able to extract information from the environment. For example, while one might expect robots equipped with proprioceptive sensors to use these alone for pose estimation, agents can also learn to utilise vision, if extracting a pose from images is not too difficult.

Finally, analogously to how \emph{data augmentation}~\cite{lecun1998gradient,shorten2019survey} can be used to improve the generalisation of models in supervised learning settings, \emph{domain randomisation} (Figure \ref{fig:dr_example}) is a common training paradigm in ``sim2real'' approaches for learning real-world robots control~\cite{andrychowicz2018learning,james2017transferring,peng2018sim,sadeghi2017cad2rl,tobin2017domain}. In DR, various properties of the simulation are varied, altering anything from the positions or dynamical properties of objects to their visual appearance, resulting in an expansion of the training domain. We therefore test how the agents trained with or without DR differ. In support of previous results, we show that agents trained with DR are more robust to out-of-distribution (OoD) perturbations. In our experiments, we use a standard visual DR setup. At each timestep, we randomise all of the components in the environment by changing/adding different textures, colour gradients, RGB colours, and noise, except for the target object. More details about the DR textures are described in Table~\ref{tbl:random_summary} and visualised in Figure~\ref{fig:random_vis}.

%\subsection{Domain Randomisation}
\begin{table}[h!]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c|c|c}
    \toprule
    No. & Option & Description \\
    \midrule
   1 & checkerboard & Render checkerboard pattern with two random RGB colours\\
   2 & gradient & Render gradient in vertical/horizontal direction with two random RGB colours\\
   3 & flat\_rgb & Render single random RGB colour\\
   4 & noise & Render artificial noise with two random RGB colours\\
    \bottomrule
  \end{tabular}}
   \caption[DR textures used during training.]{DR textures used during training. At every environment timestep, for each component (e.g., robot arm, skybox, table; the Jaco robot has 13 components, and the Fetch robot has 19 components), the appearance of the component is rendered using a sequential sampling process. In the first step of the sampling process, one of each of the 4 possible choices is made with the same probability, then the subsequent appearance of each component is determined by a second draw, described in the rightmost column of the table. Each RGB channel is drawn uniformly from $\{0 \ldots 255\}$. The details of each option are described in Table~\ref{tbl:texture_expression}.}
\label{tbl:random_summary}
\end{table}
